Assessing the Global COVID19 Impact on Air Transport with Open Data

### 2021-06-15 paper.Rmd
initital stab

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

@@

# # Purpose Statement
# 
# ???The purpose of this research is to 
# investigate additional dimensions in which 
# resilience could be measured, in addition
# to the current framework of delay analysis.
# 
# Research Question
# 
# ???How can we enlarge the concept of 
# resilience, so that it is applicable to
# scenarios of low traffic? 
```

@@

-   data-analytical approach - using open data / freely available (tbd: validated against organisational data)
-   ...

@@

@@

to be developed

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)

```

```{r}
#CURRENT COMMENTS AND TO-DO'S
# I HAVE DOWNLOADED 3 FILES FOR NOW (APR/19, APR/2020, APR/2021), JUST TO START "TIDYING" AND EXPLORING.
#IT'S IN THE DATA-RAW FOLDER (NOT SHARED WITH GITHUB), AS ALWAYS.
```

@@

# # Purpose Statement
# 
# ???The purpose of this research is to 
# investigate additional dimensions in which 
# resilience could be measured, in addition
# to the current framework of delay analysis.
# 
# Research Question
# 
# ???How can we enlarge the concept of 
# resilience, so that it is applicable to
# scenarios of low traffic? 
#
# ???Research Question: 
# ???RQ1.What was the impact of the pandemic on ATM resilience?
# ???  RQ1.1 How resilience can be modeled in a low-demand scenario?
# ???  RQ1.2 How resilient were different ATM systems worldwide?

```

@@

-   data-analytical approach - using open data / freely available (tbd: validated against organisational data)
-   ???RQ1.1 = through a qualitative analysis of previous proposed models
-   ???RQ1.2 = through a quantitative analysis of open data

@@

A mixed-method approach, based on:

  a) to answer RQ1.1, a qualitative analysis of previous models to develop acute low-demand as a disturbance
  b) to answer RQ1.2, a quantitative analysis of open data, to observe (or not) different levels/stages of stress/recovery, which could indicate different "more" or "less" resilience to the disturbances

@@

1.1

a) Resilience can be measured as a function of time - the smaller the relationship between time of stress and the time of recovery, more resilient a system is.

1.2 how to use open data to "see" resilience?

1.2.1 Gather and prepare data
```{r}
#Reading and saving data
source("./R/list_apt_files.R")
file_names <- list_apt_files()
open_sky <- map_dfr(file_names, read_csv)
#write_csv(open_sky, "./data/open_sky.csv")

# Take a look
glimpse(open_sky)
str(open_sky)
summary(open_sky)

# First exploration
fb1 <- open_sky %>% transmute(typecode = as.factor(typecode), origin = as.factor(origin), destination = as.factor(destination), firstseen, lastseen)
str(fb1)
summary(fb1)
# More than one million NA's

fb2 <- fb1 %>% drop_na()
str(fb2)
summary(fb2)
# Better...

#Visualizing something...
df1 <- fb2 %>% group_by(DATE = date(firstseen)) %>% summarize(count = n())
df1 %>% ggplot(aes(x = as.factor(year(DATE)))) +
  geom_boxplot(aes(y = count))

```

1.2.2
